### 2022-06-19 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@3b65ca139ef26a20c1cacfb5d00881f916148bae 🚀

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\emph{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.}
\textbf{Strydom2022FooWeb?}) or functional traits to fill the values of
additional species we would like to project in this space (here for
nodes I, J, K, and L) from the embedding of known species (here, nodes
A, B, C, and D). Because embeddings can be projected back to a graph,
this allows us to reconstruct a network with these new species
(\textbf{D}). This entire cycle constitutes an instance of transfer
learning, where the transfered information is the representation of
graph \textbf{A} through its embedding.}\label{fig:embedding}
}
\end{figure}

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on nodes metadata. This example has many
parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in which
node metadata can be given by phylogeny or functional traits. Rather
than directly predicting biological rules (see \emph{e.g.} Pichler et
al., 2020 for an overview), which may be confounded by the sparse nature
of graph data, learning embeddings works in the low-dimensional space
that maximizes information about the network structure. This approach is
further justified by the observation, for example, that the
macro-evolutionary history of a network is adequately represented by
some graph embeddings (RDPG; see Dalla Riva \& Stouffer, 2016). In a
recent publication, (\textbf{Strydom2022FooWeb?}) have used an embedding
(based on RDPG) to project a metaweb of trophic interactions between
European mammals, and transfered this information to mammals of Canada
by using the phylogenetic distance between related clades to infer the
values in the latent sub-space into which the metaweb was projected. By
performing the RDPG step on re-constructed value, this approach yields a
probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (\textbf{Strydom2022FooWeb?}) and t-SVD (Poisot et
al., 2021) typically embed networks using fewer dimensions than the
original network (the original network has as many dimensions as
species, and as many informative dimensions as trophically unique
species; Strydom, Dalla Riva, et al., 2021). But this is not
necessarilly the case -- indeed, one may perform a PCA (a special case
of SVD) to project the raw data into a subspace that improves the
efficacy of t-SNE (Maaten, 2009). There are many dimensionality
reductions (Anowar et al., 2021) that can be applied to an embedded
network should the need for dimensionality reduction (for example for
data visualisation) arise. In brief, many graph embeddings \emph{can}
serve as dimensionality reduction steps, but not all do, neither do all
dimensionality reduction methods provide adequate graph embedding
capacities. In the next section (and fig.~\ref{fig:illustration}), we
show how the amount of dimensionality reduction can affect the quality
of the embedding.

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using truncated
Singular Value Decomposition (t-SVD) and RDPG (see
\textbf{Strydom2022FooWeb?} for the full details). Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are
matrices that when multiplied give an approximation of the original
network. The code to reproduce this example is available as
supplementary material (note, for the sake of comparison, that Strydom,
Catchen, et al., 2021 have an example using embedding through PCA
followed by prediction using a deep neural network on the same dataset).
The resulting (binary) metaweb \(\mathcal{M}\) has 2131 interactions
between 206 parasites and 121 hosts, and its adjacency matrix has full
rank (\emph{i.e.} it represents a space with 121 dimensions). All
analyses were done using Julia (Bezanson et al., 2017) version 1.7.2,
\emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

@@

>>>>

\begin{figure}
\hypertarget{fig:embedding}{%
\centering
\includegraphics{figures/conceptual_2.png}
\caption{Overview of the embedding process. A network (\emph{A}),
represented here as its adjacency matrix, is converted into a
lower-dimensional object (\emph{B}) where nodes, subgraphs, or edges
have specific values (see tbl.~\ref{tbl:methods} for an overview of
methods and their use for species interactions). For the purposes of
prediction, this low-dimensional object encodes feature vectors for
\emph{e.g.} the nodes. Embedding also allows to visualize the structure
in the data differently (see fig.~\ref{fig:illustration}), much like
with a principal component analysis. From a low-dimensional feature
vector, it is possible to develop predictive approaches. Nodes in an
ecological network are usually species (\textbf{C}), for which we can
leverage phylogenetic relatedness (\emph{e.g.} Strydom et al., 2022) or
functional traits to fill the values of additional species we would like
to project in this space (here for nodes I, J, K, and L) from the
embedding of known species (here, nodes A, B, C, and D). Because
embeddings can be projected back to a graph, this allows us to
reconstruct a network with these new species (\textbf{D}). This entire
cycle constitutes an instance of transfer learning, where the transfered
information is the representation of graph \textbf{A} through its
embedding.}\label{fig:embedding}
}
\end{figure}

@@

The popularity of graph embedding techniques in machine learning is more
than the search for structural invariants: graphs are discrete objects,
and machine learning techniques tend to handle continuous data better.
Bringing a sparse graph into a continuous, dense vector space (Xu, 2020)
opens up a broader variety of predictive algorithms, notably of the sort
that are able to predict events as probabilities (Murphy, 2022).
Furthermore, the projection of the graph itself is a representation that
can be learned; Runghen et al. (2021), for example, used a neural
network to learn the embedding of a network in which not all
interactions were known, based on nodes metadata. This example has many
parallels in ecology (see fig.~\ref{fig:embedding} \textbf{C}), in which
node metadata can be given by phylogeny or functional traits. Rather
than directly predicting biological rules (see \emph{e.g.} Pichler et
al., 2020 for an overview), which may be confounded by the sparse nature
of graph data, learning embeddings works in the low-dimensional space
that maximizes information about the network structure. This approach is
further justified by the observation, for example, that the
macro-evolutionary history of a network is adequately represented by
some graph embeddings (RDPG; see Dalla Riva \& Stouffer, 2016). In a
recent publication, Strydom et al. (2022) have used an embedding (based
on RDPG) to project a metaweb of trophic interactions between European
mammals, and transfered this information to mammals of Canada by using
the phylogenetic distance between related clades to infer the values in
the latent sub-space into which the metaweb was projected. By performing
the RDPG step on re-constructed value, this approach yields a
probabilistic trophic metaweb for mammals of Canada based on knowledge
of European species, despite a limited (\(\approx\) 5\%) taxonomic
overlap.

Graph embeddings \emph{can} serve as a dimensionality reduction method.
For example, RDPG (Strydom et al., 2022) and t-SVD (Poisot et al., 2021)
typically embed networks using fewer dimensions than the original
network (the original network has as many dimensions as species, and as
many informative dimensions as trophically unique species; Strydom,
Dalla Riva, et al., 2021). But this is not necessarilly the case --
indeed, one may perform a PCA (a special case of SVD) to project the raw
data into a subspace that improves the efficacy of t-SNE (Maaten, 2009).
There are many dimensionality reductions (Anowar et al., 2021) that can
be applied to an embedded network should the need for dimensionality
reduction (for example for data visualisation) arise. In brief, many
graph embeddings \emph{can} serve as dimensionality reduction steps, but
not all do, neither do all dimensionality reduction methods provide
adequate graph embedding capacities. In the next section (and
fig.~\ref{fig:illustration}), we show how the amount of dimensionality
reduction can affect the quality of the embedding.

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using truncated
Singular Value Decomposition (t-SVD) and RDPG (see Strydom et al., 2022
for the full details). Briefly, an RDPG decomposes a network into two
subspaces (left and right), which are matrices that when multiplied give
an approximation of the original network. The code to reproduce this
example is available as supplementary material (note, for the sake of
comparison, that Strydom, Catchen, et al., 2021 have an example using
embedding through PCA followed by prediction using a deep neural network
on the same dataset). The resulting (binary) metaweb \(\mathcal{M}\) has
2131 interactions between 206 parasites and 121 hosts, and its adjacency
matrix has full rank (\emph{i.e.} it represents a space with 121
dimensions). All analyses were done using Julia (Bezanson et al., 2017)
version 1.7.2, \emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

@@

\leavevmode\hypertarget{ref-Strydom2022FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L. J., Runghen,
R., Dalla Riva, G. V., \& Poisot, T. (2022). Food web reconstruction
through phylogenetic transfer of low-rank network representation.
\emph{Methods in Ecology and Evolution}, \emph{n/a}(n/a).
\url{https://doi.org/10.1111/2041-210X.13835}
